---
title: Are college teams passing more on first down?
author: Shane Connuck
date: '2022-11-10'
slug: are-college-teams-passing-more-on-first-down
categories: []
tags: []
---


```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(cfbfastR)
library(ggalt)
```

```{r}
tictoc::tic()
pbp <- data.frame()
seasons <- 2014:cfbfastR:::most_recent_cfb_season()
progressr::with_progress({
  
  pbp <- cfbfastR::load_cfb_pbp(seasons)
})
tictoc::toc()
```
The following code creates a new dataframe for all passing plays on first down, grouped by the team and year.
```{r}
pass_plays <- pbp %>% 
  filter(pass == 1) %>% 
  filter(down == 1) %>% 
  group_by(pos_team, year) %>%
  summarise(plays = n())
```
This a new dataframe containing all first down plays. 
```{r}
plays <- pbp %>% 
  filter(down == 1) %>% 
  group_by(pos_team, year) %>%
  summarise(plays = n())
```
So, I want to join these two new dataframes together so I can see the first down passes and plays side-by-side.
```{r}
plays <- pass_plays %>% left_join(plays, by=c("pos_team", "year"))
```
Now, let's get the percentage of first down plays that were passes:
```{r}
pct_plays <- plays %>% 
  group_by(year, plays.x, plays.y) %>% 
  summarise(first_pass_pct = plays.x/plays.y)
```
```{r}
summary(pct_plays)
```

On average, about 32 percent of plays on first down are passes. Now, let's make this specific for each year.
```{r}
pct_plays %>% 
  group_by(first_pass_pct, year) %>% 
  summary(year == "2021")
```


```{r}
ggplot() + 
  geom_line(data=plays, aes(x=pos, y=year)) + 
  scale_y_continuous(limits = c(0, 500))
```

